A molecular dynamics study of the solvation of carbon dioxide and other compounds in the ionic liquids \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}

### 2018-09-15 ionic_liquids.tex
Some changes in Dilute CO2 solvation

@@

If both $P_s$ and $c_s$ are small, we can resort to Eqs.~\eqref{eq:solvation free energy definition} and \eqref{eq:free_solv} to predict Henry's law constant, which is defined as $\text{K}_s =\frac{P_s}{x_s}$ \cite{Prausnitz}.
%\begin{equation}
%\text{K}_s =\frac{P_s}{x_s}.
%\end{equation}
%
Therefore, by simply expressing Eq.~\eqref{eq:solvation free energy definition} in terms of the molar fraction of the solute in the liquid phase ($x_s$) and assuming a dilute solution, whose molar density does not perceivably differ from that of pure solvent (that is, $\tilde{\rho} \approx \tilde{\rho}_\text{\tiny IL}$), we obtain that
\begin{equation}
\label{eq:henry_eq}
K_s = \tilde{\rho}_\text{\tiny IL} R \, T \exp\left(\frac{\Delta G_\text{sim}}{k_b T}\right).
\end{equation}

@@

\subsection{Henry's Law Constants of \ce{CO_2} in \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}}
\label{sec:henry_results}

In Table \ref{table:henry} we present the results of the solvation free energies of \ce{CO_2} along with the Henry constants ($K_s$) obtained by means of Eq.~\eqref{eq:henry_eq}.
Table~\ref{table:henry} also contains the experimental $K_s$ reported by Mahurin \textit{et al}. \cite{Mahurin_2010} and Finotello \textit{et al}. \cite{Finotello_2008}.
The uncertainties in $K_s$ were calculated using standard error propagation.

\begin{table*}
	\centering
	\caption{Results of solvation free energies and Henry constants of \ce{CO_2} in \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]} at $T = 298.15~\mathrm{K}$ and $P = 1~\mathrm{atm}$. The experimental Henry constants of \ce{CO_2} in \ce{[emim][B(CN)_4]} and in \ce{[emim][NTf_2]} are, respectively, $38.9 \pm 0.03 ~\mathrm{atm}$ \cite{Mahurin_2010} and $39.0 \pm 0.1 ~\mathrm{atm}$ \cite{Finotello_2008}.}
	\begin{adjustbox}{max width=\textwidth}
		\begin{tabular}{ccccccc}  
			\hline\hline
			Model & $\tilde{\rho}_\text{sim}$ & $\Delta G_\text{\tiny vdW}$  & $\Delta G_\text{\tiny Coul}$  & $\Delta G_\text{sim}$ & $K_s$ & error \\
			& (mol/dm\textsuperscript{3}) & (kcal/mol) & (kcal/mol) &  (kcal/mol) & (atm)  & (\%) \\
			\hline
%			\multicolumn{7}{c}{\ce{[emim][B(CN)_4]}: $K_{s}^{\text{exp}} = 38.9 \pm 0.03 ~\mathrm{atm}$ \cite{Mahurin_2010}} \\
			\multicolumn{7}{c}{\ce{[emim][B(CN)_4]}} \\
			Koller \textit{et al}. \cite{Koller_2012} & $4.5925$ & $0.45 \pm 0.01$ & $-1.06 \pm 0.02$ & $-0.61 \pm 0.02$ & $40 \pm 1$ & $2.8$ \\
			Batista \textit{et al}. \cite{Batista_2015} & $4.3700$ & $0.378 \pm 0.008$ & $-1.24 \pm 0.01$  & $-0.86 \pm 0.01$ & $24.9 \pm 0.5$ & $38.9$ \\
			Liu \textit{et al}. \cite{Liu_2014} & $4.3815$ & $0.226 \pm 0.009$ & $-1.11 \pm 0.01$ & $-0.88 \pm 0.01$ & $24.3 \pm 0.5$ & $37.6$  \\
			Weber and Kirchner \cite{Weber_2016} & $4.5221$ & $0.478 \pm 0.009$ & $-1.23 \pm 0.02$ & $-0.75 \pm 0.02$ & $31 \pm 1$ & $20.3$  \\
			\hline
%			\multicolumn{7}{c}{\ce{[emim][NTf_2]}: $K_{s}^{\text{exp}} = 39.0 \pm 0.1 ~\mathrm{atm}$ \cite{Finotello_2008} }\\
			\multicolumn{7}{c}{\ce{[emim][NTf_2]}} \\
			K\"{o}ddermann \textit{et al}. \cite{K_ddermann_2007} & $3.8205$ & $0.389 \pm 0.005$ & $-0.901 \pm 0.006$ & $-0.512 \pm 0.008$ & $39.4 \pm 0.5$  & $0.97$  \\
			\hline\hline
			\label{table:henry} 
		\end{tabular}
	\end{adjustbox}
\end{table*}

It should be pointed out that the result of $\Delta G_\text{sim}$ obtained for the model of Liu \textit{et al}. \cite{Liu_2014} is statistically identical to the value reported by those authors \cite{Liu_2014_1}, who applied the BAR method \cite{Bennett_1976} to independent sampling of $\lambda_{\,\text{vdW}}$ and $\lambda_{\, \text{Coul}}$.
Moreover, the $\Delta G_\text{sim}$ in \ce{[emim][NTf_2]} is very close to the value obtained by Kerl \textit{et al}. \cite{Kerl__2017} ($-0.534$ kcal/mol, with no uncertainty reported) using the FEP method \cite{Zwanzig_1954}.
These consistencies have served for validating the methodological approaches of this work.

From Table \ref{table:henry} one can see that the model of Koller \textit{et al}. \cite{Koller_2012} best reproduces the experimental value of $K_s$.
If we look at the $\Delta G_{\lambda_{\,\text{vdW}}}$ profiles in Fig.~\ref{fig:deltag}, we observe that for the models of Batista \textit{et al}. \cite{Batista_2015} and Liu \textit{et al}.~\cite{Liu_2014} the maximum occurs at lower values of energy, suggesting that these models underestimate the interatomic interactions, which may also explain their failure to reproduce the experimental density.

\begin{figure}
\includegraphics[width=\linewidth]{free_energy_paper}
\caption{$\Delta G_{\lambda_{\,\text{vdW}}}$ profiles corresponding to the insertion of a \ce{CO_2} molecule in \ce{[emim][B(CN)_4]}, which was modeled according to the force fields of (circles) Koller \textit{et al}. \cite{Koller_2012}, (squares) Batista \textit{et al}. \cite{Batista_2015}, (up triangles) Liu \textit{et al}. \cite{Liu_2014}, and (x) Weber and Kirchner \cite{Weber_2016}.}
\label{fig:deltag}
\end{figure}

The $\Delta G_{\, \text{sim}}$ values for the models of Koller \textit{et al}. \cite{Koller_2012} and K\"{o}ddermann \textit{et al}. \cite{K_ddermann_2007} suggest a more favorable solvation of \ce{CO_2} in \ce{[emim][B(CN)_4]} than in \ce{[emim][NTf_2]}. 
The spatial distribution functions (SDFs) \cite{Svishchev_1993} shown in Figs.~\ref{fig:sdf_ions} confirm this result.
The SDFs of the ions around a \ce{CO_2} molecule were obtained by using the software packages TRAVIS \cite{Brehm_2011} and VMD \cite{HUMP96}.
For that, we carried out NPT-MD simulations of \ce{CO_2}-\ce{[emim][B(CN)_4]} and \ce{CO_2}-\ce{[emim][NTf_2]} systems with a \ce{CO_2} molar fraction of 0.1.
The isovalues were visually determined so as to observe the first solvation shell.
If we compare Figs.~\ref{fig:sdf_ions} (a) and (b), one notes that the region with a low probability of finding an ion is larger for \ce{[emim][NTf_2]}.
Also, the first solvation shell due to [B(CN)\textsubscript{4}]$^{-}$ occurs at a slightly shorter distance than for \ce{[NTf_2]^-}, as one sees from Fig.~\ref{fig:sdf_ions} (c).

@@

\subsection{Study of concentrated solutions of \ce{CO_2} in \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}}
\label{sec:results_conc}

>>>>

\usepackage{xcolor}
\usepackage{pdfcomment}

@@

If both $P_s$ and $c_s$ are small, we can resort to Eqs.~\eqref{eq:solvation free energy definition} and \eqref{eq:free_solv} to predict Henry's law constant, which is defined as $\text{K}_s =\frac{P_s}{x_s}$ \cite{Prausnitz}.
%\begin{equation}
%\text{K}_s =\frac{P_s}{x_s}.
%\end{equation}
%
Therefore, by simply expressing Eq.~\eqref{eq:solvation free energy definition} in terms of the molar fraction of the solute in the liquid phase ($x_s$), we can conclude that
\begin{equation}
\label{eq:henry_eq}
K_s = \tilde{\rho}_\text{sim} R \, T \exp\left(\frac{\Delta G_\text{sim}}{k_b T}\right),
\end{equation}
where $\tilde{\rho}_\text{sim}$ is the simulated molar density of the solution. Assuming that this solution is very dilute, one can use the molar density of pure solvent ($\tilde{\rho}_\text{\tiny IL}$) instead.

@@

\subsection{Dilute Solutions of \ce{CO_2} in \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}}
\label{sec:henry_results}

In Table \ref{table:henry}, we present results obtained from expanded ensemble simulations of a single \ce{CO_2} molecule dissolved in 250 ion pairs of either \ce{[emim][B(CN)_4]} or \ce{[emim][NTf_2]} under periodic boundary conditions.
The simulations were carried out at $T = 298.0~\mathrm{K}$ and $P = 1.0~\mathrm{bar}$.
Table~\ref{table:henry} contains simulated molar densities, partial and total solvation free energies, as well as Henry's law constants ($K_s$) computed via Eq.~\eqref{eq:henry_eq}.
It also contains the experimental $K_s$ values reported by Mahurin \textit{et al}. \cite{Mahurin_2010} and by Finotello \textit{et al}. \cite{Finotello_2008}.
The uncertainties in $K_s$ were calculated by using a standard error propagation formula.

\begin{table*}
	\centering
	\caption{Results of solvation free energies and Henry's law constants of \ce{CO_2} in \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]} at $T = 298.15~\mathrm{K}$ and $P = 1~\mathrm{atm}$. The experimental Henry constants of \ce{CO_2} in \ce{[emim][B(CN)_4]} and in \ce{[emim][NTf_2]} are, respectively, $38.9 \pm 0.03 ~\mathrm{atm}$ \cite{Mahurin_2010} and $39.0 \pm 0.1 ~\mathrm{atm}$ \cite{Finotello_2008}.}
	\begin{adjustbox}{max width=\textwidth}
		\begin{tabular}{ccccccc}  
			\hline\hline
			Model & $\tilde{\rho}_\text{sim}$ & $\Delta G_\text{\tiny vdW}$  & $\Delta G_\text{\tiny Coul}$  & $\Delta G_\text{sim}$ & $K_s$ & error \\
			& (mol/dm\textsuperscript{3}) & (kcal/mol) & (kcal/mol) &  (kcal/mol) & (atm)  & (\%) \\
			\hline
%			\multicolumn{7}{c}{\ce{[emim][B(CN)_4]}: $K_{s}^{\text{exp}} = 38.9 \pm 0.03 ~\mathrm{atm}$ \cite{Mahurin_2010}} \\
			\multicolumn{7}{c}{\ce{[emim][B(CN)_4]}} \\
			Koller \textit{et al}. \cite{Koller_2012} & $4.5925$ & $0.45 \pm 0.01$ & $-1.06 \pm 0.02$ & $-0.61 \pm 0.02$ & $40 \pm 1$ & $2.8$ \\
			Batista \textit{et al}. \cite{Batista_2015} & $4.3700$ & $0.378 \pm 0.008$ & $-1.24 \pm 0.01$  & $-0.86 \pm 0.01$ & $24.9 \pm 0.5$ & $38.9$ \\
			Liu \textit{et al}. \cite{Liu_2014} & $4.3815$ & $0.226 \pm 0.009$ & $-1.11 \pm 0.01$ & $-0.88 \pm 0.01$ & $24.3 \pm 0.5$ & $37.6$  \\
			Weber and Kirchner \cite{Weber_2016} & $4.5221$ & $0.478 \pm 0.009$ & $-1.23 \pm 0.02$ & $-0.75 \pm 0.02$ & $31 \pm 1$ & $20.3$  \\
			\hline
%			\multicolumn{7}{c}{\ce{[emim][NTf_2]}: $K_{s}^{\text{exp}} = 39.0 \pm 0.1 ~\mathrm{atm}$ \cite{Finotello_2008} }\\
			\multicolumn{7}{c}{\ce{[emim][NTf_2]}} \\
			K\"{o}ddermann \textit{et al}. \cite{K_ddermann_2007} & $3.8205$ & $0.389 \pm 0.005$ & $-0.901 \pm 0.006$ & $-0.512 \pm 0.008$ & $39.4 \pm 0.5$  & $0.97$  \\
			\hline\hline
			\label{table:henry} 
		\end{tabular}
	\end{adjustbox}
\end{table*}

The result of $\Delta G_\text{sim}$ obtained with the model of Liu \textit{et al}. \cite{Liu_2014} is statistically identical to the value reported by the same authors \cite{Liu_2014_1}, who applied the Bennett Acceptance Ratio (BAR) method \cite{Bennett_1976} to independent samples obtained with different values of $\lambda_\text{\tiny vdW}$ and $\lambda_\text{\tiny Coul}$.
Moreover, the $\Delta G_\text{sim}$ result for \ce{[emim][NTf_2]} is very close to the value obtained by Kerl \textit{et al}. \cite{Kerl__2017} ($-0.534 ~\mathrm{kcal/mol}$, with no uncertainty reported) using the Free-Energy Perturbation (FEP) method \cite{Zwanzig_1954}.
%These consistencies have served for validating the methodological approaches of this work.

From Table \ref{table:henry}, one can see that no model other than that of Koller \textit{et al}. \cite{Koller_2012} was able to reproduce the experimental Henry's law constant of \ce{CO_2} in \ce{[emim][B(CN)_4]} with small deviation.
If we look at the $\Delta G_\text{\tiny vdW}$ profiles in Fig.~\ref{fig:deltag}, we observe that for the models of Batista \textit{et al}. \cite{Batista_2015} and Liu \textit{et al}.~\cite{Liu_2014} the maximum occurs at lower values of energy, suggesting that these models underestimate the interatomic interactions, which may also explain their failure to reproduce the experimental density.

\begin{figure}
\includegraphics[width=\linewidth]{free_energy_paper}
\caption{Profile of $\Delta G_\text{\tiny vdW}$ corresponding to the insertion of a \ce{CO_2} molecule in \ce{[emim][B(CN)_4]}, modeled according to the force fields of Koller \textit{et al}. \cite{Koller_2012} ($\bullet$), Batista \textit{et al}. \cite{Batista_2015} ($\blacksquare$), Liu \textit{et al}. \cite{Liu_2014} ($\times$), and Weber and Kirchner \cite{Weber_2016} ($\blacktriangle$).}
\label{fig:deltag}
\end{figure}

The $\Delta G_{\, \text{sim}}$ values for the models of Koller \textit{et al}. \cite{Koller_2012} and K\"{o}ddermann \textit{et al}. \cite{K_ddermann_2007} suggest a more favorable solvation of \ce{CO_2} in \ce{[emim][B(CN)_4]} than in \ce{[emim][NTf_2]}. 
The spatial distribution functions (SDFs) \cite{Svishchev_1993} shown in Figs.~\ref{fig:sdf_ions} confirm this result.
The SDFs of the ions around a \ce{CO_2} molecule were obtained by using the software packages TRAVIS \cite{Brehm_2011} and VMD \cite{HUMP96}.
For that, we carried out NPT-MD simulations of \ce{CO_2}-\ce{[emim][B(CN)_4]} and \ce{CO_2}-\ce{[emim][NTf_2]} systems with a \ce{CO_2} molar fraction of 0.1.
The isovalues were visually determined so as to observe the first solvation shell.
If we compare Figs.~\ref{fig:sdf_ions} (a) and (b), one notes that the region with a low probability of finding an ion is larger for \ce{[emim][NTf_2]}.
Also, the first solvation shell due to \ce{[B(CN)_4^-} occurs at a slightly shorter distance than for \ce{[NTf_2]^-}, as one sees from Fig.~\ref{fig:sdf_ions}(c).

@@

\subsection{Concentrated solutions of \ce{CO_2} in \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}}
\label{sec:results_conc}
